---
title: "Reproducible Research Course Project 1"
author: "LI Shaobai"
date: "2020/6/22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

READING DATA
------------
```{r reading_data}
dat=read.csv('activity.csv')
```

TOTAL STEPS PER DAY
-------------------
```{r total_steps_per_day}
s1=split(dat,dat$date)
totalsteps=as.data.frame(sapply(s1,function(x) sum(x$steps)))   #calculate sums
names(totalsteps)='steps'
print(totalsteps)       #show the total number of steps taken per day
hist(totalsteps$steps, main='Histogram of Daily Total Steps', xlab='steps')     #plot histogram
summary(totalsteps)
```
**As the data summary shows, the mean and the median of the total number of steps taken per day are  10766 and 10765 respectively.**

AVERAGE DAILY ACTIVITY PATTERN
------------------------------
```{r average_daily_activity_pattern}
s2=split(dat,dat$interval)
stepsper5min=as.data.frame(sapply(s2,function(x) mean(x$steps,na.rm=TRUE)))     #drop NAs for means calculation
names(stepsper5min)='steps'
stepsper5min$interval=row.names(stepsper5min)
with(stepsper5min,plot(interval,steps,type='l',main='Average Steps per 5 min')) #making the time series plot
stepsper5min$interval[which.max(stepsper5min$steps)]     #search for the maximum 5-minute interval
```
**So the 5-minute interval of 8:35-8:40 contains the maximum number of steps.**

IMPUTING MISSING VALUES
-----------------------
```{r imputing_missing_values}
summary(dat)    #check NAs in the original data set
#creating a new data set by imputing missing steps with the mean for that 5-minute interval
newdat=dat
for (i in 1:nrow(newdat)) {
     if(is.na(newdat$steps[i])==TRUE){
         index=which(stepsper5min$interval==newdat$interval[i])
         newdat$steps[i]=stepsper5min$steps[index]
     }
 }
s3=split(newdat,newdat$date)
newtotalsteps=as.data.frame(sapply(s3,function(x) sum(x$steps)))        #calculate sums
names(newtotalsteps)='steps'
hist(newtotalsteps$steps, main='New Histogram of Daily Total Steps', xlab='steps')      #plot histogram
summary(newtotalsteps)
```
**As the summary of the original data set shows, there are 2304 NAs only in the 'steps' column, i.e. there are 2304 rows with NAs totally;**

**My imputing strategy is to use the mean steps for that 5-minute interval;**

**As the summary of the new data set shows, the mean and the median of the total number of steps taken per day are both 10766, of which the median is 1 more than the original while the mean is the same;**

**As the two histograms shows, imputing missing data with the mean for that 5_minute interval gathers more data towards the middle of the plot.**

DIFFERENCE BETWEEN WEEKDAYS AND WEEKENDS
----------------------------------------
```{r difference_between_weekdays_and_weekends}
for (j in 1:nrow(newdat)) {
     if(weekdays(as.Date(newdat$date[j]))=='星期六'|weekdays(as.Date(newdat$date[j]))=='星期日'){
         newdat$day[j]='weekend'
     }
     else{newdat$day[j]='weekday'}
 }
newdat$day=as.factor(newdat$day)        #create the new factor variable
str(newdat)     #show the new factor variable
#calculate for weekdays and weekends respectively
weekdaydat=subset(newdat,day=='weekday')
s4=split(weekdaydat,weekdaydat$interval)
weekdaysteps=as.data.frame(sapply(s4,function(x) mean(x$steps,na.rm=TRUE)))
names(weekdaysteps)='steps'
weekdaysteps$interval=row.names(weekdaysteps)
weekdaysteps$day=rep('weekday',nrow(weekdaysteps))
weekenddat=subset(newdat,day=='weekend')
s5=split(weekenddat,weekenddat$interval)
weekendsteps=as.data.frame(sapply(s5,function(x) mean(x$steps,na.rm=TRUE)))
names(weekendsteps)='steps'
weekendsteps$interval=row.names(weekendsteps)
weekendsteps$day=rep('weekend',nrow(weekendsteps))
steps=rbind(weekdaysteps,weekendsteps)  #combine the two dataframe together for ggplot
library(ggplot2)
ggplot(steps,aes(interval,steps,group=1))+geom_line()+facet_grid(day~.)+theme(axis.text.x  = element_blank())        #hide x scale for clarity
```

